---
title: "RepData_PeerAssessment1"
author: "Kristophe"
date: "9/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading the csv file to be used for analysis.

```{r cars}
#Checking the working directory
getwd()

library(readr)
library(data.table)
library(dplyr)
library(knitr)
library(ggplot2)
library(ggthemes)
library(Hmisc)

#Loading the data
activity=read.csv("activity/activity.csv")

# Viewing the dataset
View(activity)

```

Checking the structure of the data and observing some of the data

```{r, echo=TRUE}
# Observing the first 10 observations
head(activity,10)

# Observing the last 10 observations
tail(activity,10)

#Checking the structure of the data
str(activity)

# Changing the date format
activity$date=as.Date(activity$date)

str(activity)

```

Getting the total number of steps

```{r}
# Getting the total number of steps
sum(activity$steps,na.rm=TRUE)

```

Plotting a histogram of the total number of steps taken per day
```{r, fig.height=13,fig.width=13}
hist(activity$steps,main = "Histogram of Total Number of Steps per Day",xlab = "Total Steps per day")

```

Getting the total mean and median of the total number of steps per day

```{r}
mean(activity$steps,na.rm = TRUE)

median(activity$steps,na.rm = TRUE)

```


Making a time series plot of 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r,fig.width=13,fig.height=13}
plot(activity$interval,activity$steps,xlab="Number of Steps taken",ylab = "5-Minutes Interval",main="Time series Plot",type = "l")
```

Getting the number of missing values and replacing with the mean

```{r}
colSums(is.na(activity))

for(i in 1:ncol(activity)){
  activity[is.na(activity[,i]), i] <- mean(activity[,i], na.rm = TRUE)
}

colSums(is.na(activity))
```

Plotting a Histogram and Getting the mean and median

```{r}
hist(activity$steps,main = "Histogram of Total Number of Steps per Day",xlab = "Total Steps per day")

```

Checking if there are differences in activity patterns between weekdays and weekends

```{r}

activity$date <- ifelse(weekdays(activity$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
```

Plotting a time series plot

```{r}

xyplot(activity$steps~activity$interval|activity$date,data=activity,type="l")

```